---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EndangeRed

<!-- badges: start -->
<!-- badges: end -->

The goal of EndangeRed is to ...

## Installation

You can install the development version of EndangeRed like so:

``` r
remotes::install_github("https://github.com/CENTRO-INTERNACIONAL-DE-LA-PAPA/EndangeRed")
```

## Example

This is a basic example which shows you how to get the Red Listing of varieties.

First, we need to calculate the Overall Cultivar Frequency (OCF) and Relative Cultivar Frequency (RCF).

```{r example}
library(EndangeRed)

data(varieties_data)

ocf_data <- OCF(
    dfr = varieties_data,
    vname = "variety_name",
    hh = "household_code",
    community = "community",
    location = "location"
)

print(head(ocf_data))

rcf_data <- RCF(
    dfr = varieties_data,
    vname = "variety_name",
    hh = "household_code",
    nsvarie = "number_of_tubers",
    community =
        "community",
    location = "location"
)

print(head(rcf_data))

```

Now, we can plot the OCF and RCF values in normal scale:

```{r, message=FALSE}

Plot_Variable_Red_Listing(ocf_data, rcf_data,type = "normal")

```

Or in log scale:

```{r, message=FALSE}

Plot_Variable_Red_Listing(ocf_data, rcf_data,type = "log")

```


If we need to get the varieties inside each quadrant we can do it by using the function `Get_Red_Listing`.

```{r}

combined_scales <- Get_Red_Listing(ocf_data, rcf_data,type = "log")

print(combined_scales)

red_listing_varieties <- combined_scales$varieties[[1]]

print(red_listing_varieties)

```
