---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EndangeRed

<!-- badges: start -->
<!-- badges: end -->

The goal of EndangeRed is to provide a tool to calculate the Red Listing of varieties based on the Overall Cultivar Frequency (OCF) and Relative Cultivar Frequency (RCF) values.

## Installation

You can install the development version of EndangeRed like this:

``` r
remotes::install_github("https://github.com/CENTRO-INTERNACIONAL-DE-LA-PAPA/EndangeRed")
```

## Example

This is a basic example which shows you how to get the Red Listing of varieties based on the Overall Cultivar Frequency (OCF) and Relative Cultivar Frequency (RCF) values.

Here we have two years of data, 2013 and 2017 on the same location.

```{r example}

library(EndangeRed)
library(dplyr)
library(tidyr)
data(varieties_data)

table(varieties_data$Año)

```
First let's select a year. For this example we will use 2013.

```{r}

varieties_data_2013 <- varieties_data %>%
    filter(Año == 2013)

ocf_data <- OCF(
    dfr = varieties_data_2013,
    vname = "variety_name",
    hh = "household_code",
    community = "community",
    location = "location"
)

print(head(ocf_data))

rcf_data <- RCF(
    dfr = varieties_data_2013,
    vname = "variety_name",
    hh = "household_code",
    nsvarie = "number_of_tubers",
    community =
        "community",
    location = "location"
)

print(head(rcf_data))

```

Now, we can plot the OCF and RCF values in normal scale by using the function `Plot_Variable_Red_Listing`.

```{r, message=FALSE}

Plot_Variable_Red_Listing(ocf_data, rcf_data,type = "normal")

```

Or in log scale:

```{r, message=FALSE}

Plot_Variable_Red_Listing(ocf_data, rcf_data,type = "log")

```


If we need to get the varieties inside each quadrant we can do it by using the function `Get_Red_Listing`.

```{r}

combined_scales <- Get_Red_Listing(ocf_data, rcf_data,type = "log")

print(combined_scales)

red_listing_varieties <- combined_scales$varieties[[1]]

print(red_listing_varieties)

```

Now, what if we want to get the Red Listing of both years? We can do it by using the following code:

```{r}

library(patchwork)
library(purrr)

nested_data <- varieties_data %>% 
    group_by(Año) %>% 
    nest()

nested_data <- nested_data %>% 
    mutate(
        OCF = map(data, ~OCF(
            dfr = .x,
            vname = "variety_name",
            hh = "household_code",
            community = "community",
            location = "location"
        )),
        RCF = map(data, ~RCF(
            dfr = .x,
            vname = "variety_name",
            hh = "household_code",
            nsvarie = "number_of_tubers",
            community = "community",
            location = "location"
        )),
        Plots = pmap(list(OCF, RCF,Año), ~Plot_Variable_Red_Listing(.x, .y,type = "log") + 
                         patchwork::plot_annotation(
                             subtitle = Año,
                             theme = ggplot2::theme(plot.subtitle = ggplot2::element_text(size = 12, hjust = 0.5, colour = "black"))
                         )),
        combined_scales = map2(OCF, RCF, ~Get_Red_Listing(.x, .y,type = "log"))
    )

nested_data

```

Notice that we have created a list of plots and a list of combined scales for each year. We can access the data by using the following code:

For the year 2013:

```{r}
nested_data$Plots[[1]]
nested_data$combined_scales[[1]]$varieties[[1]]
```

For the year 2017:

```{r}
nested_data$Plots[[2]]
nested_data$combined_scales[[2]]$varieties[[1]]
```

